Continuous developmental change can explain discontinuities in word learning

### 2019-02-01 cogsci/cogsci.Rmd
final version cogsci

Nevertheless, at least in some cases, development may only appear to be stage-like. This appearance can be due, for example, to the use of a cognitively-demanding task which may mask learning, or to the use of statistical thresholding (in particular, p-value < 0.05) which can create a spurious dichotomy between success and failure in observing a given behavior. In such cases, positing discontinuous stages is unnecessary. Instead, a continuous model–involving similar representations across the lifespan–may provide a simpler and more transparent account of development.

We use a case study from word learning literature. @stager1997 first showed that children's early ability to distinguish similar sounds does not automatically translate into word learning skills. Indeed, though infants around 14-month old can distinguish similar sound pairs such as "dih" and "bih", they appear to fail in mapping this pair to two different objects. Subsequent research has focused on proposing possible explanations for this observed gap between speech perception and word learning [e.g., @fennell2010; @stager1997; @hofer2017;  @rost2009]. In this work we focus, rather, on the mechanism of development.

@@

Subsequent findings have suggested otherwise. On the one hand, 14-month-olds–who typically fail in the original task–succeed when an easier testing method is used, even under the same learning conditions [@yoshida2009]. They also succeed when uncertainty is mitigated via disambiguating cues [e.g., @thiessen2007]. On the other hand, adults show patterns of learning similar to those shown by 14-month-olds when the task is more challenging and when similarity between words increases [@white2013; @pajak2016].

@@

\begin{SCfigure}
\centering
\includegraphics[width=1.5in]{figs/model.png}
\caption{Graphical representation of our model. Circles indicate random variables (shading indicates observed variables). The squares indicates fixed model parameters.}
\label{fig:model}
\end{SCfigure}

A similar probabilistic structure was used by @lewis2013 to model concept learning, and by @hofer2017 to model spoken word learning. However, the first study assumed that the sounds are heard unambiguously, and the second assumed the concepts are observed unambiguously. In our model, we assume that both labels and concepts are observed with a certain amount of perceptual noise, which we assume, for simplicity, that it is captured by a normal distribution:

@@

## Inference
<!--If we denote the correct/target object as $o_T$ and the incorrect object as $o_{\lnot T}$, then the probability of chosing the correct object is: $P(o_T|s_T) = $ 
(see Figure \ref{fig:task})
-->
The learner hears a sound $s$ and has to decide which object $o$ provides an optimal match to this sound. To this end, the learner has to compute the probability $P(o|s)$ for all possible objects. This probability can be computed by summing over all possible concepts and labels:
$$P(o|s)=\sum_{C,L} P(o, C, L| s) \propto \sum_{C,L} P(o, C, L, s) $$
The joint probability $P(o, C, L, s)$ is obtained by factoring the Bayesian network in Figure \ref{fig:model}:
$$P(o,C,L,s) = P(s|L)P(L|C)P(C|o)P(o) $$
which could be transformed using Bayes rule into:

$$P(o,C,L,s) = P(s|L)P(L|C)P(o|C)P(C) $$
Finally, assuming that the concepts' prior probability is uniformly distributed\footnote{This is a reasonable assumption given the similarity of the concepts used in each naming situation. See the stimuli sub-section in Experiment.}, we obtain the following expression, where all conditional dependencies are now well defined:

@@

We use the model to predict performance in the word learning task introduced by @stager1997, and a testing method similar to the one used by @yoshida2009. In this task, participants are first exposed to the association between pairs of nonsense words (e.g., "lif"/"neem") and pairs of objects. The word-object associations are introduced sequentially. After this exposure phase, participants perform a series of test trials. In each of these trials, one of the two sounds is uttered (e.g., “lif”) and participants choose the corresponding object from the two alternatives. An overview of the task is shown in Figure \ref{fig:task}.

We use the general Equation 1 to derive the exact analytical expression for the probability of accurate responses $p(o_T | s)$ (target object $o_T$ given a sound $s$) in the simple case of two-alternative forced choice in the testing phase of our experimental task:

@@

In this experiment, we tested participants in the word learning task introduced above (Figure \ref{fig:task}). More precisely, we explored the predictions related to both distinctiveness and precision. Sound similarity ($\Delta s$) and object similarity ($\Delta o$) were varied simultaneously in a within-subject design. Two age groups (preschool children and adults) were tested on the same task to explore whether development can be characterized with the uncertainty parameters, $\sigma_C$ and $\sigma_L$. The Experiment, sample size, exclusion criteria and the model's main predictions were pre-registered.
<!--Children were trained to learn the association between pairs of nonesense words (e.g., "lif"/"neem") and pairs of objects (e.g., two flowers). After training, children performed a series of two-alternative forced choices. In each testing trial, one of the two sounds is uttered (e.g., “lif”) and participants choose the corresponding object from the two alternatives (e.g., flower 1 or flower 2). Each child learned an artificial lexicon where both the pairs of sounds and the pairs of objects can be similar or different. Our predictions were as follows. If the sound and meaning interact in word learning, then the pair of words that differ on both the meaning and sound levels should be the easiest to learn, followed by the pairs of words that are different on one level, but similar on another level. Finally, the pair of words that are similar on both levels should be the most difficult to learn.  -->

@@

### Participants
We planned to recruit a sample of N=60 children ages 4-5 years from the Bing Nursery School on Stanford University’s campus. So far, we have collected data from n=`r N_good_ch` children. An additional n=`r N_all_ch-N_good_ch` children participated but were removed from analyses because they were not above chance on the catch trials. We also planned to recruit a sample of N=60 adults on Amazon Mechanical Turk. This study used sata from n=`r N_good_ad` adult participants. An additional n=`r N_all_ad-N_good_ad` participants were tested but excluded because of low scores on the catch trials (n=`r N_all_ad-N_good_ad-d_adult_lang` ) or because they were familiar with the non-English sound stimuli we used in the adult experiment (n=`r d_adult_lang`).

@@

In a separate survey, $N=20$ participants recruited on Amazon Mechanical Turk evaluated the similarity of each sound and object pair on a 7-point scale. We scaled responses within the range [0,1]. Data are shown in Figure \ref{fig:stim}, for each stimuli group. This data will be used in the models as the perceptual distance of sound pairs ($\Delta s$) and object pairs ($\Delta o$).

@@

###Design 
Each age group saw only two of the three levels of similarity described in the previous sub-section: "different" vs. "intermediate" for preschoolers, and "intermediate" vs. "similar" for adults. We made this choice in the light of pilot studies showing that adults were at ceiling with "different" sounds/objects, and children were at chance with the "similar" sounds/objects. That said, this difference in the level of similarity is accounted for in the model through using the appropriate perceptual distance used in each age group (Figure \ref{fig:stim}).

```{r stim, fig.env = "figure", fig.pos = "h", fig.align='center', fig.width=3.5, fig.height=2.5, fig.cap = "\\label{fig:stim}Distances for both sound and object pairs from an adult norming study. Error bars represent 95\\% confidence intervals."}

sim_plot
```
The experiment was a 2x2 within-subjects factorial design with four conditions: high/low sound similarity crossed with high/low visual object similarity. Besides the 4 conditions, we also tested participants on a fifth catch condition which was similar in its structure to the other ones, but was used only to select participants who were able to follow the instructions and show minimal learning.

@@

#plot_data <- ggplot(subset(data_human_model, source =='Experimental'), 
plot_data <- ggplot(data_human_model, 
      aes(x = sound_dist, y = value, group = factor(concept_dist), col = factor(concept_dist))) +
  #geom_point()+
  geom_pointrange(aes(ymin = ci.lower, ymax = ci.upper), 
                 position = position_dodge(width = .1),
                 size=0.2) + 
  geom_line() + 
  #geom_line(data=subset(data_human_model, source == "Model"), linetype=2, size=0.5, col='black')+
  #geom_point(data=subset(data_human_model, source == "Model"), size=1, col='black')+
  geom_hline(yintercept = 0.50, linetype="dashed") +
  xlab("Auditory similary") +ylab("Accuracy")+
  #scale_colour_discrete(name="Visual similarity")+
  theme_few()+
  ggthemes::scale_color_ptol(name="Visual similarity")+
theme(legend.title = element_text(size=11),
      legend.text=element_text(size=11),
      axis.text = element_text(size = 11)
      #) + facet_grid(group~source, scales="free_y") +
      ) + facet_grid(source~group) +
  theme(aspect.ratio = 0.7)

@@

Preschoolers were tested at the nursery school using a tablet, whereas adults used their own computers to complete the same experiment online. Participants were tested in a sequence of five conditions: the four experimental conditions plus the catch condition. In each condition, participants saw a first block of four exposure trials followed by four testing trials, and a second block of two exposure trials (for memory refreshment) followed by an additional four testing trials.

```{r all_data, fig.env = "figure*", fig.pos = "h", fig.align = "center", fig.width=6, fig.height=4, fig.cap = "\\label{fig:data_all}Accuracy of novel word recognition as as a function of the sound distance, the object distance, and the age group (preschool children vs. Adults). We show both experimental results (top) and model predictions (bottom). Error bars represent 95\\% confidence intervals."}
plot_data
#img <- png::readPNG("figs/data_all.png")
#grid::grid.raster(img)

@@

Experimental results are shown in Figure 6 (top). We first analyzed the results using a mixed-effects logistic regression with sound distance, object distance and age group as fixed effects, and with a maximal random effects structure (allowing us to take into account the full nested structure of our data) [@barr2013]. We found main effects for all the fixed effects in the regression, i.e., the sound distance ($\beta =$ `r s_coef`, $p$ `r p_val(s_p)`)–replicating previous findings, the object distance ($\beta =$ `r c_coef`, $p$ `r p_val(c_p)`)–confirming the new prediction of our model, and age group ($\beta =$ `r age_coef`, $p$ `r p_val(age_p)`)–showing that performance improves with age. Besides, we found two-way interactions between sound distance and age ($\beta =$ `r s_age_coef`, $p$ `r p_val(s_age_p)`) and between object distance and age ($\beta =$ `r c_age_coef`, $p$ `r p_val(c_age_p)`).

@@

Second, we fit our model (using Equation 2) to the participants' responses in each age group.  The values of $\Delta s$ and $\Delta o$ were set based on data from the similarity judgment task (Figure \ref{fig:stim}). Thus, the model has two degrees of freedom for each group, i.e., $\sigma_C$ and $\sigma_L$.  Figure \ref{fig:data_all} (bottom) shows the predictions. The model captures the qualitative patterns in both age groups: starting from a low accuracy recognition when both the sound and object distances are small, the model correctly predicts an increase in accuracy when either the sound distance or the object distance increases. Besides, accuracy is correctly predicted to be maximal when both the sound and object distances are high.

The values of the parameters were as follows. Children had a label-specific uncertainty of $\sigma_S =$ `r round(mod2_val$val[which(mod2_val$group=="children" & mod2_val$param=="Vr_s")],2)` [`r round(mod2_val$lower[which(mod2_val$group=="children" & mod2_val$param=="Vr_s")],2)`, `r round(mod2_val$upper[which(mod2_val$group=="children" & mod2_val$param=="Vr_s")],2)`]\footnote{All uncertainty intervals in this paper represent 95\% Confidence Intervals.}, and a concept-specific uncertainty of $\sigma_C =$ `r round(mod2_val$val[which(mod2_val$group=="children" & mod2_val$param=="Vr_o")],2)` [`r round(mod2_val$lower[which(mod2_val$group=="children" & mod2_val$param=="Vr_o")],2)`, `r round(mod2_val$upper[which(mod2_val$group=="children" & mod2_val$param=="Vr_o")],2)`]. Adults had a label-specific uncertainty of $\sigma_S =$ `r round(mod2_val$val[which(mod2_val$group=="adults" & mod2_val$param=="Vr_s")],2)` [`r round(mod2_val$lower[which(mod2_val$group=="adults" & mod2_val$param=="Vr_s")],2)`, `r round(mod2_val$upper[which(mod2_val$group=="adults" & mod2_val$param=="Vr_s")],2)`], and a concept-specific uncertainty of $\sigma_C =$ `r round(mod2_val$val[which(mod2_val$group=="adults" & mod2_val$param=="Vr_o")],2)` [`r round(mod2_val$lower[which(mod2_val$group=="adults" & mod2_val$param=="Vr_o")],2)`, `r round(mod2_val$upper[which(mod2_val$group=="adults" & mod2_val$param=="Vr_o")],2)`].  As predicted, the uncertainty parameters were larger for children than they were for adults, showing that the probabilsitic representations becomes more refined (that is, $\sigma$ becomes smaller) across development.

The models explained the majority of the variance in the participants' mean responses ($R^2=$ `r R2_model2`, for the combined adult and children's data). To investigate whether the model's predictive power is due to overfitting, we fit a simplified version with only one degree of freedom (i.e., a single variance common to both sounds and objects). This single-variance model explained as much variance in the mean responses ($R^2=$ `r R2_model1`). It also captured the main qualitative patterns (graph not shown), suggesting that the explanatory power of the model is largely due to its structure, rather than its degrees of freedom.

An unexpected outcome was that adult participants deviated slightly from the model's numbers: While the model predicted  accuracy to be more sensitive to object distance when sound distance is higher (and vice-versa), adult participants showed the opposite pattern. This deviation is interesting as it suggests that participants are more likely to pay attention and integrate additional sources of information when ambiguity is higher. \footnote{It is unclear whether the same pattern is found in children as the difference between accuracy scores is generally small compared to the uncertainty of the measurements.}

@@

This paper explored the idea that some seemingly stage-like patterns in cognitive development can be characterized in a continuous fashion. We used as a case study the seminal work of @stager1997 showing a discrepancy between children's speech perception abilities and their word learning skills. While most previous work debated the source of this discrepancy, here we were more interested in how it reflects developmental change.

<!--Since the seminal work of @stager1997 showing a discrepancy between children's speech perception abilities and their word learning skills, researchers have debated about the source of such a discrepancy. In this work, we were more interested in the development of this phenomenon, regardless of its exact origin. -->

@@

The model made a new prediction to which we provided the first experimental test: Learning similar words is not only modulated by the similarity of their phonological forms, but also by the visual similarity of their semantic referents. More generally, since visual similarity is an early organizing feature in the semantic domain [e.g.,@wojcik2013], our finding suggests that children may prioritize the acquisition of words that are quite distant in the semantic space.  This suggestion is supported by recent findings based on the investigation of early vocabulary growth [@engelthaler2017; @sizemore2018].

@@

One limitation of this work is that the model was fit to data from children at a relatively older age (4-5 years old) than what is typically studied in the literature (around 14 month-old). Nonetheless, this choice was related to two measurement difficulties. First, model fitting requires several, reliably-measured data points. In our case, data collection involved presenting participants with several trials across four conditions in a between-subject design. It would have been challenging to obtain these measures with infants. Second, the study of infants requires using specific testing methods. This fact would have made the quantitative study of development challenging. Indeed, it is not clear how data from these methods would directly compare to data collected using more typical methods in older children and adults.
<!--While an important conclusion of this paper is that the same process may apply to different developmental stages, it would be interesting to experimentally verify the generality of this conclusion in the context of early language acquisition. A challenge to this next step would be to construct a computational framework which would compare performance across heterogeneous testing methods (e.g., looking time vs. forced-choice). -->

In sum, this paper proposes a model that accounts for the development of an important aspect of word learning. The findings show that data can be explained based on a continuous process operating over similar representations across development. We used a case from word learning as an example, but the same idea might apply to other aspects of cognitive development that are typically thought of as stage-like (e.g., acquisition of a theory of mind). Computational models such as the one proposed here can help us investigate the extent to which such discontinuities emerge due to genuine qualitative changes, and the extent to which they reflect the granularity of the researchers' own measurement tools.

>>>>

Nevertheless, at least in some cases, development may only appear to be stage-like. This appearance can be due, for example, to the use of a cognitively-demanding task which may mask learning, or to the use of statistical thresholding (in particular, p-value < 0.05) which can create a spurious dichotomy between success and failure in observing a given behavior. In such cases, positing discontinuous stages is unnecessary. Instead, a continuous model---involving similar representations across the lifespan---may provide a simpler and more transparent account of development.

We use a case study from word learning literature. @stager1997 first showed that children's early ability to distinguish similar sounds does not automatically translate into word learning skills. Indeed, though infants around 14-month old can distinguish similar sound pairs such as "dih" and "bih", they appear to fail in mapping this pair to two different objects. Follow-up studies has focused on proposing possible explanations for this observed gap between speech perception and word learning [e.g., @fennell2010; @stager1997; @hofer2017;  @rost2009].

<!--In this work we focus, rather, on the mechanism of development. -->

@@

Subsequent findings have suggested otherwise. On the one hand, 14-month-olds---who typically fail in the original tas---succeed when an easier testing method is used, even under the same learning conditions [@yoshida2009]. They also succeed when uncertainty is mitigated via disambiguating cues [e.g., @thiessen2007]. On the other hand, adults show patterns of learning similar to those shown by 14-month-olds when the task is more challenging and when similarity between words increases [@white2013; @pajak2016].

@@

\begin{SCfigure}
\centering
\includegraphics[width=1.5in]{figs/model.png}
\caption{Graphical representation of our model. Circles indicate random variables (shading indicates observed variables). The squares indicate fixed model parameters.}
\label{fig:model}
\end{SCfigure}

A similar probabilistic structure was used by @lewis2013 to model concept learning, and by @hofer2017 to model spoken word learning. However, the first study assumed that the sounds are heard unambiguously, and the second assumed the concepts are observed unambiguously. In our model, we assume that both labels and concepts are observed with a certain amount of perceptual noise, which we assume, for simplicity, is captured by a normal distribution:

@@

## Inference
<!--If we denote the correct/target object as $o_T$ and the incorrect object as $o_{\lnot T}$, then the probability of chosing the correct object is: $P(o_T|s_T) = $ 
(see Figure \ref{fig:task})
-->
The learner hears a sound $s$ and has to decide which object $o$ provides an optimal match to this sound. To this end, they must compute the probability $P(o|s)$ for all possible objects. This probability can be computed by summing over all possible concepts and labels:
$$P(o|s)=\sum_{C,L} P(o, C, L| s) \propto \sum_{C,L} P(o, C, L, s) $$
The joint probability $P(o, C, L, s)$ is obtained by factoring the Bayesian network in Figure \ref{fig:model}:
$$P(o,C,L,s) = P(s|L)P(L|C)P(C|o)P(o) $$
which can be transformed using Bayes rule into:

$$P(o,C,L,s) = P(s|L)P(L|C)P(o|C)P(C) $$
Finally, assuming that the concepts' prior probability is uniformly distributed\footnote{This is a reasonable assumption in our particular case given the similarity of the concepts used in each naming situation in our experiment.}, we obtain the following expression, where all conditional dependencies are now well defined:

@@

  We use the model to predict performance in the word learning task introduced by @stager1997, with a two-alternative forced choice as in @yoshida2009. In this task, participants are first exposed to the association between pairs of nonsense words (e.g., "lif"/"neem") and pairs of objects. The word-object associations are introduced sequentially. After this exposure phase, participants perform a series of test trials. In each of these trials, one of the two sounds is uttered (e.g., “lif”) and participants choose the corresponding object from the two alternatives. An overview of the task is shown in Figure \ref{fig:task}.

We used Equation 1 to obtain the exact analytical expression for the probability of accurate responses $p(o_T | s)$ (target object $o_T$ given a sound $s$) in the simple case of two-alternative forced choice in the testing phase of our experimental task:

@@

In this experiment, we tested participants in the word learning task introduced above (Figure \ref{fig:task}). More precisely, we explored the predictions related to both distinctiveness and precision. Sound similarity ($\Delta s$) and object similarity ($\Delta o$) were varied simultaneously in a within-subject design. Two age groups (preschool children and adults) were tested on the same task to explore whether development can be characterized with the uncertainty parameters, $\sigma_C$ and $\sigma_L$. The experiment, sample size, exclusion criteria and the model's main predictions were pre-registered.
<!--Children were trained to learn the association between pairs of nonesense words (e.g., "lif"/"neem") and pairs of objects (e.g., two flowers). After training, children performed a series of two-alternative forced choices. In each testing trial, one of the two sounds is uttered (e.g., “lif”) and participants choose the corresponding object from the two alternatives (e.g., flower 1 or flower 2). Each child learned an artificial lexicon where both the pairs of sounds and the pairs of objects can be similar or different. Our predictions were as follows. If the sound and meaning interact in word learning, then the pair of words that differ on both the meaning and sound levels should be the easiest to learn, followed by the pairs of words that are different on one level, but similar on another level. Finally, the pair of words that are similar on both levels should be the most difficult to learn.  -->

@@

### Participants
We planned to recruit a sample of N=60 children ages 4-5 years from the Bing Nursery School on Stanford University’s campus. Here we report data from n=`r N_good_ch` children. An additional n=`r N_all_ch-N_good_ch` children participated but were removed from analyses because they were not above chance on the catch trials due to the challenging nature of our procedure (see below). We also planned to recruit a sample of N=60 adults on Amazon Mechanical Turk. Data from n=`r N_all_ad-N_good_ad` participants were excluded due to low scores on the catch trials (n=`r N_all_ad-N_good_ad-d_adult_lang` ) or because they were familiar with the non-English sound stimuli we used in the adult experiment (n=`r d_adult_lang`), yielding a final sample of n=`r N_good_ad`.

@@

In a separate survey, $N=20$ participants recruited on Amazon Mechanical Turk evaluated the similarity of each sound and object pair on a 7-point scale. We scaled responses within the range [0,1]. Data are shown in Figure \ref{fig:stim}, for each stimulus group. These data will be used in the models as the perceptual distance of sound pairs ($\Delta s$) and object pairs ($\Delta o$).

@@

###Design

Each age group saw only two of the three levels of similarity described in the previous sub-section: "different" vs. "intermediate" for preschoolers, and "intermediate" vs. "similar" for adults. We made this choice in light of pilot studies showing that adults were at ceiling with "different" sounds/objects, and children were at chance with the "similar" sounds/objects. That said, this difference in the level of similarity is accounted for in the model through using the appropriate perceptual distance used in each age group (Figure \ref{fig:stim}).

```{r stim, fig.env = "figure", fig.pos = "h", fig.align='center', fig.width=3.5, fig.height=2.5, fig.cap = "\\label{fig:stim}Distances for both sound and object pairs from an adult norming study. Data represent Likert values normalized to [0,1] interval. Error bars represent 95\\% confidence intervals."}

sim_plot
```

To maximize our ability to measure subtle stimulus effects, the experiment was a 2x2 within-subjects factorial design with four conditions: high/low sound similarity crossed with high/low visual object similarity. Besides the 4 conditions, we also tested participants on a fifth catch condition which was similar in its structure to the other ones, but was used only to select participants who were able to follow the instructions and show minimal learning.

@@

plot_data <- ggplot(subset(data_human_model, source =='Experimental'), 
#plot_data <- ggplot(data_human_model, 
      aes(x = sound_dist, y = value, group = factor(concept_dist), col = factor(concept_dist))) +
  #geom_point()+
  geom_pointrange(aes(ymin = ci.lower, ymax = ci.upper), 
                 position = position_dodge(width = .1),
                 size=0.2) + 
  geom_line() + 
  geom_line(data=subset(data_human_model, source == "Model"), linetype=2, size=0.5)+
  geom_point(data=subset(data_human_model, source == "Model"), size=1.5, alpha=0.5)+
  geom_hline(yintercept = 0.50, linetype="dashed") +
  xlab("Auditory similary") +ylab("Accuracy")+
  #scale_colour_discrete(name="Visual similarity")+
  theme_few()+
  ggthemes::scale_color_ptol(name="Visual similarity")+
theme(legend.title = element_text(size=11),
      legend.text=element_text(size=11),
      axis.text = element_text(size = 11)
      #) + facet_grid(group~source, scales="free_y") +
      ) + facet_grid(.~group) +
  theme(aspect.ratio = 0.7)

@@

Preschoolers were tested at the nursery school using a tablet, whereas adults used their own computers to complete the same experiment online. Participants were tested in a sequence of five conditions: the four experimental conditions plus the catch condition. In each condition, participants saw a first block of four exposure trials followed by four testing trials, and a second block of two exposure trials (for memory refreshment) followed by an additional four testing trials. The length of this procedure was demanding, especially for children, but we adopted a fully within-subjects design based on pilot testing that indicated that precision of measurement was critical for testing our experimental predictions.

```{r all_data, fig.env = "figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "\\label{fig:data_all}Accuracy of novel word recognition as as a function of the sound distance, the object distance, and the age group (preschool children vs. Adults). We show both experimental results (solid lines) and model predictions (dashed lines). Error bars represent 95\\% confidence intervals."}
plot_data
#img <- png::readPNG("figs/data_all.png")
#grid::grid.raster(img)

@@

Experimental results are shown in Figure 6 (top). We first analyzed the results using a mixed-effects logistic regression with sound distance, object distance and age group as fixed effects, and with a maximal random effects structure (allowing us to take into account the full nested structure of our data) [@barr2013]. We found main effects for all the fixed effects in the regression. For the sound distance, we obtained $\beta =$ `r s_coef` ($p$ `r p_val(s_p)`), replicating previous findings. For object distance, we found $\beta =$ `r c_coef`  ($p$ `r p_val(c_p)`), and this finding confirms the new prediction of our model. Finally, for the age group, we obtained  $\beta =$ `r age_coef` ($p$ `r p_val(age_p)`), showing that performance improves with age. In addition, we found two-way interactions between sound distance and age ($\beta =$ `r s_age_coef`, $p$ `r p_val(s_age_p)`) and between object distance and age ($\beta =$ `r c_age_coef`, $p$ `r p_val(c_age_p)`).

@@

We next fit our model (using Equation 2) to the participants' responses in each age group.  The values of $\Delta s$ and $\Delta o$ were set based on data from the similarity judgment task (Figure \ref{fig:stim}). Thus, the model has two degrees of freedom for each group, i.e., $\sigma_C$ and $\sigma_L$.  Figure \ref{fig:data_all} (bottom) shows the predictions. The model captures the qualitative patterns in both age groups: starting from a low accuracy recognition when both the sound and object distances are small, the model correctly predicts an increase in accuracy when either the sound distance or the object distance increases. Further, accuracy is correctly predicted to be maximal when both the sound and object distances are high.

The values of the parameters were as follows. Children had a label-specific uncertainty of $\sigma_S =$ `r round(mod2_val$val[which(mod2_val$group=="children" & mod2_val$param=="Vr_s")],2)` [`r round(mod2_val$lower[which(mod2_val$group=="children" & mod2_val$param=="Vr_s")],2)`, `r round(mod2_val$upper[which(mod2_val$group=="children" & mod2_val$param=="Vr_s")],2)`]\footnote{All uncertainty intervals in this paper represent 95\% Confidence Intervals.}, and a concept-specific uncertainty of $\sigma_C =$ `r round(mod2_val$val[which(mod2_val$group=="children" & mod2_val$param=="Vr_o")],2)` [`r round(mod2_val$lower[which(mod2_val$group=="children" & mod2_val$param=="Vr_o")],2)`, `r round(mod2_val$upper[which(mod2_val$group=="children" & mod2_val$param=="Vr_o")],2)`]. Adults had a label-specific uncertainty of $\sigma_S =$ `r round(mod2_val$val[which(mod2_val$group=="adults" & mod2_val$param=="Vr_s")],2)` [`r round(mod2_val$lower[which(mod2_val$group=="adults" & mod2_val$param=="Vr_s")],2)`, `r round(mod2_val$upper[which(mod2_val$group=="adults" & mod2_val$param=="Vr_s")],2)`], and a concept-specific uncertainty of $\sigma_C =$ `r round(mod2_val$val[which(mod2_val$group=="adults" & mod2_val$param=="Vr_o")],2)` [`r round(mod2_val$lower[which(mod2_val$group=="adults" & mod2_val$param=="Vr_o")],2)`, `r round(mod2_val$upper[which(mod2_val$group=="adults" & mod2_val$param=="Vr_o")],2)`].  As predicted, the uncertainty parameters were larger for children than they were for adults, showing that the probabilistic representations becomes more refined (that is, $\sigma$ becomes smaller) across development. The developmental effect was more important for the label-specific uncertainty.

The models explained the majority of the variance in the participants' mean responses ($R^2=$ `r R2_model2`, for the combined adult and children's data). To investigate whether the model's predictive power was due to overfitting, we fit a simplified version with only one degree of freedom (i.e., a single variance common to both sounds and objects). This single-variance model explained as much variance in the mean responses ($R^2=$ `r R2_model1`). It also captured the main qualitative patterns (graph not shown), suggesting that the explanatory power of the model is largely due to its structure, rather than its degrees of freedom.

An unexpected outcome was that adult participants deviated slightly from the model's numbers: While the model predicted  accuracy to be more sensitive to object distance when sound distance is higher (and vice-versa), adult participants showed the opposite pattern. This deviation is an interesting starting point for future work as it may suggest that participants are more likely to pay attention to and integrate additional sources of information when ambiguity is higher.

@@

This paper explored the idea that some seemingly stage-like patterns in cognitive development can be characterized in a continuous fashion. We used as a case study the seminal work of @stager1997 showing a discrepancy between children's speech perception abilities and their word learning skills. While much of the previous investigation of this finding has been interested in the source of this discrepancy, here we have explored how it could arise from continuous developmental change in perceptual uncertainty.
<!--Since the seminal work of @stager1997 showing a discrepancy between children's speech perception abilities and their word learning skills, researchers have debated about the source of such a discrepancy. In this work, we were more interested in the development of this phenomenon, regardless of its exact origin. -->

@@

The model made a new prediction to which which we tested experimentally: Learning similar words is not only modulated by the similarity of their phonological forms, but also by the visual similarity of their semantic referents. More generally, since visual similarity is an early organizing feature in the semantic domain [e.g.,@wojcik2013], our finding suggests that children may prioritize the acquisition of words that are quite distant in the semantic space.  This suggestion is supported by recent findings based on the investigation of early vocabulary growth [@engelthaler2017; @sizemore2018].

@@

One limitation of this work is that the model was fit to data from children at a relatively older age (4-5 years old) than what is typically studied in the literature (14-18 month-old). We selected this age group based on two measurement issues. First, model fitting requires several, reliably-measured data points. In our case, data collection involved presenting participants with several trials across four conditions in a between-subject design. It would have been challenging to obtain these measures with infants. Second, experiments with infants on this topic have often used habituation methods [but cf. @yoshida2009], which require additional linking assumptions. Indeed, it is not clear how data from these methods would directly compare to data collected using more typical methods in older children and adults.
<!--While an important conclusion of this paper is that the same process may apply to different developmental stages, it would be interesting to experimentally verify the generality of this conclusion in the context of early language acquisition. A challenge to this next step would be to construct a computational framework which would compare performance across heterogeneous testing methods (e.g., looking time vs. forced-choice). -->

In sum, this paper proposes a model that accounts for the development of an important aspect of word learning. Our account suggests that the developmental data can be explained based on a continuous process operating over similar representations across development, suggesting developmental continuity. We used a case from word learning as an example, but the same idea might apply to other aspects of cognitive development that are typically thought of as stage-like (e.g., acquisition of a theory of mind). Computational models such as the one proposed here can help us investigate the extent to which such discontinuities emerge due to genuine qualitative changes, and the extent to which they reflect the granularity of the researchers' own measurement tools.
